---
title: "Reproducible Research Project 1"
author: "HsuanJu Chen"
date: "2018/11/05"
output: html_document
---

## Loading  package
```{r message=FALSE}
library(dplyr)
library(ggplot2)
```

## Load the data
```{r}
Sys.setlocale("LC_TIME", "English")

Dataframe <- read.csv("activity.csv", 
                      sep = ",",
                      header = TRUE,
                      stringsAsFactors = FALSE)
```

## Transform the data 
```{r}
Dataframe$steps <- as.numeric(Dataframe$steps)
Dataframe$date <- as.Date(Dataframe$date, "%Y-%m-%d")
SubDataframe <- na.omit(Dataframe) # Remove missing values 
```


## Q1: mean total number of steps taken per day
### Calculate the total number of steps taken per day
```{r}
(stepsPerDay <- SubDataframe %>%
    group_by(date) %>%
    summarize(sum(steps)))
```

### Make a histogram of the total number of steps taken each day
```{r}
hist(stepsPerDay$`sum(steps)`,
     xlab = "Total Number of Steps Per Day", 
     ylab = "Frequency",
     main = "Histogram of total number of steps per day",
     breaks = 10, 
     xlim = c(0, 25000), 
     ylim = c(0, 20),
     col = "blue")
```


### Calculate and report the mean and median of the total number of steps taken per day
```{r}
(mean(stepsPerDay$`sum(steps)`))
(median(stepsPerDay$`sum(steps)`))
```
#### The mean total number of steps taken per day is 10766.19 and the median is 10765


## Q2: What is the average daily activity pattern?
### Make a time series plot 
```{r}
(stepsPerInterval <- SubDataframe %>% 
    group_by(interval) %>%
    summarize(mean(steps)))
```

```{r}
plot(stepsPerInterval$interval,stepsPerInterval$`mean(steps)`,
     type = "l", 
     xlab = "Time Interval (5-mins)", 
     ylab = "Average number of steps",
     main = "Average Number of Steps by Time  Interval",
     col = "blue")
```

### Find the interval with the maximum number of steps
```{r}
(maxInterval<- stepsPerInterval$interval[which.max(stepsPerInterval$`mean(steps)`)])
```
#### The maximum number of steps was observed during interval 835


## Q3: Imputing missing values

### Calculate and report the total number of missing values in the dataset
```{r}
(missingValues <-sum (is.na(Dataframe$steps)))
```
#### There are 2304 missing values in the dataset

### Create a new dataset without the missing data 
```{r}
newDataframe <- Dataframe
for(i in 1:nrow(newDataframe)){
    if(is.na(newDataframe$steps[i])){
        newDataframe$steps[i] <- mean(newDataframe$steps, na.rm = TRUE)
    }
}
```

### Plot the histogram of steps per day for the dataset without the missing values 
```{r}
(newstepsPerDay <- newDataframe %>%
    group_by(date) %>%
    summarize(sum(steps)))
```

```{r}
hist(newstepsPerDay$`sum(steps)`,
     xlab = "Total Number of Steps Per Day", 
     ylab = "Frequency",
     main = "Histogram of total number of steps per day",
     breaks = 10, 
     xlim = c(0, 25000), 
     ylim = c(0, 20),
     col = "blue")
```

### Calculate and report the mean and median total number of steps taken per day.
```{r}
(mean(newstepsPerDay$`sum(steps)`))
(median(newstepsPerDay$`sum(steps)`))
```
#### The mean total number of steps taken per day is 10766.19 and the median is 10766.19


## Q4: Are there differences in activity patterns between weekdays and weekends?
### Create a variable with weekdays 
```{r}
newDataframe$day <- weekdays(newDataframe$date)

# Define days that are Saturday or Sunday as weekends
newDataframe <- newDataframe %>%
    mutate(dayClass = ifelse(day %in% c("Saturday", "Sunday"),"weekend",
                             "weekday")) 
```

### Make a panel plot containing a time series plot 
```{r}
(newstepsPerDay2 <- newDataframe %>%
    group_by(dayClass, interval) %>%
    summarize(mean(steps)))
```

```{r}
ggplot(newstepsPerDay2, aes(x = interval, y = `mean(steps)`)) + 
    geom_line(color = "blue") + 
    xlab("Time Interval (5-mins)") + 
    ylab("Average number of steps") + 
    ggtitle("Average Number of Steps by Time Interval") + 
    facet_grid(dayClass~.)
```
